---
title: "Energy Consumption Analysis: G7 vs BRICS"
author: "Jeremy Chee"
format: html
editor: visual
---

## Project Overview

Energy consumption is closely linked to economic development, industrial activity, and population growth. Comparing energy use across country groups can provide insight into differences in development patterns and resource demand.\
\
This project analyses energy consumption per person (measured in kilowatt-hours, kWh) across two major country groups:\
\
- \*\*G7 countries:\*\* Canada, France, Germany, Italy, Japan, United Kingdom, United States\
- \*\*BRICS countries:\*\* Brazil, Russia, India, China, South Africa\
\
The goal of this analysis is to compare long-term trends in energy consumption and evaluate differences in average per-capita energy use between these groups over time.

## Data Preparation

```{r}
library(gt)
library(patchwork)
library(tidyverse)

energy_data <- read_csv("data/energy-use-per-person-kwh.csv")

str(energy_data)
colSums(is.na(energy_data))
```

## Group Definitions

```{r}
g7_countries <- c(
  "Canada", "France", "Germany", "Italy",
  "Japan", "United Kingdom", "United States"
)

brics_countries <- c(
  "Brazil", "Russia", "India", "China", "South Africa"
)

energy_data <- energy_data |>
  mutate(
    Group = case_when(
      Country %in% g7_countries ~ "G7",
      Country %in% brics_countries ~ "BRICS",
      TRUE ~ NA_character_
    )
  )

energy_subset <- energy_data |>
  filter(!is.na(Group))
```

## Analysis

```{r}
group_avg_energy <- energy_subset |>
  group_by(Year, Group) |>
  summarise(
    avg_energy_kwh = mean(Energy, na.rm = TRUE),
    .groups = "drop"
  )

group_avg_energy |>
  slice_head(n = 10) |>
  gt()
```

## Visualisation

```{r}
#| fig-width: 14
#| fig-asp: 1.2
#| fig-align: center

plot_brics <- energy_subset |>
  filter(Group == "BRICS") |>
  ggplot(aes(x = Year, y = Energy, color = Country)) +
  geom_line() +
  labs(
    title = "BRICS Countries: Energy Use per Person",
    x = "Year",
    y = "kWh per Person"
  ) +
  theme_minimal()

plot_g7 <- energy_subset |>
  filter(Group == "G7") |>
  ggplot(aes(x = Year, y = Energy, color = Country)) +
  geom_line() +
  labs(
    title = "G7 Countries: Energy Use per Person",
    x = "Year",
    y = "kWh per Person"
  ) +
  theme_minimal()

plot_avg <- group_avg_energy |>
  ggplot(aes(x = Year, y = avg_energy_kwh, color = Group)) +
  geom_line(linewidth = 1.2) +
  labs(
    title = "Average Energy Use per Person: G7 vs BRICS",
    x = "Year",
    y = "Average kWh"
  ) +
  theme_minimal()

plot_brics / plot_g7 / plot_avg
```

## Key Findings

Energy use per person is consistently higher in G7 countries than in BRICS countries across the years shown.\
\
The BRICS countries generally start from a much lower base, although some countries show clear growth over time, especially as industrial activity and development increase.\
\
The group-average comparison highlights a persistent gap between the two groups. Overall, the results suggest that G7 countries have maintained substantially higher per-capita energy consumption than BRICS countries over the period analysed.
